---
title: "An example of how to use `bbsAssistant` to download and munge the BBS data from the FTP server or local file."
author: "Jessica Burnett"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r  setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(include=FALSE, warning = FALSE, message=FALSE)
```

## Define and/or create local directories
This function will create, if it does not already exist, folder **./bbsData/** within which to locally store BBS data and results. 
**NOTE**: If the directory exists, it will not overwrite files. If the bbs data already exists inside bbsDir, then we will create a logical to NOT download the data (see below). If you wish to download more, or overwrite existing data, please specify downloadBBSData=TRUE or remove .zip files from **/bbsData/**.
```{r  createDirs}
# Create a directory to store and/or load the BBS data as feathers
bbsDir <- here::here("bbsData/")
dir.create(bbsDir)
```

# Download the BBS data, import, and save R objects to file.
If necessary, download all the state data. This takes 10-15 minutes, so only run if you have not recently downloaded the BBS data. 
```{r getFloridaData, echo=TRUE}
# a. Load the regional .txt file from Patuxent's FTP server (you must be connected to the internet to perform this step)
regions <- get_regions()

# b. Create a series or one filenames for states, regions
regionFileName <- regions$zipFileName %>% na.omit()

# c. Let's limit the data downloaded, since downloading all regions is expensive. The state or province names are truncated, so in many cases we don't want to specify eg. 'Florida.zip'
regionFileName.use <- regionFileName[which(str_detect(regionFileName, "Flori")==TRUE)]

# d.  Download and unzip the BBS raw count data to a temporary location and import as R object
flBBS <- get_bbsData(file=regionFileName.use)

# e. Get the route information
routes <- get_routeInfo()

# e. Append route information to the count data
flBBS <- left_join(flBBS, routes)
glimpse(flBBS)

# f. Save the unzipped files to disk.
export_bbsFeathers(dataIn = flBBS,
                newDir  = bbsDir,
                filename = regionFileName.use)

```

If the BBS data was downloaded previously and saved as .feather, we can import it using `import_bbsFeathers`. The code below is particularly useful if you are importing multiple files...
```{r import_bbsFeathers, echo=TRUE}
feathers <- NULL
featherNames <- list.files(bbsDir, pattern = ".feather")
featherNames <- str_c("/", featherNames) #add separator
for (i in 1:length(featherNames)) {
  feather <- NULL
  feather <- import_bbsFeathers(newDir  = bbsDir,
                              filename = featherNames[i])
  feathers <- rbind(feathers, feather)
  rm(feather)
}

# you will notice the imported data slightly differs from the original bbs data in that # cols is far fewer
ncol(feathers); ncol(flBBS)

# for now, let's just stick with the flBBS data frame. 
rm(feathers) # remove feathers from file
```

If you wish to download and/or import ALL the data, you might choose to do so in a loop. Note: this is expensive! The following are not run in this example.
```{r downloadAllBBS, eval = FALSE, echo=TRUE}
# Throw a warning if files exist
    if(length(list.files(bbsDir, pattern = "*.feather")) > 0 ){
        downloadBBSData = FALSE
    }else(
        {dir.create(bbsDir)
        downloadBBSData = TRUE}
        )
## Download ALL the regions of BBS data
if(downloadBBSData==TRUE){
for(i in 1:length(regionFileName)){
        bbsData <-  import_bbsData(
            # arguments for get_bbsData()
            file = regionFileName[i],
            dir =  "ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/States/",
            year = NULL,
            aou = NULL,
            countrynum = NULL,
            states = NULL,
            #  arguments for get_routeInfo():
            routesFile = "routes.zip",
            routesDir =  "ftp://ftpext.usgs.gov/pub/er/md/laurel/BBS/DataFiles/",
            RouteTypeID = 1,
            # one or more of c(1,2,3)
            Stratum = NULL,
            BCR = NULL
        )
# d. Save the unzipped files to disk.
export_bbsFeathers(dataIn  = bbsData,
                newDir  = bbsDir,
                filename = regionFileName[i])
# e. Clear object from memory
rm(bbsData)
} # end section I. loop
}else(message(paste0("NOT DOWNLOADING BBS DATA. If you wish to download the BBS data, please remove files from directory: ",bbsDir))) # end if-else to download the data
```
Or import a lot of data files.
```{r loadAllData, eval = FALSE}
feathers <- NULL
featherNames <- list.files(bbsDir, pattern = ".feather")
featherNames <- str_c("/", featherNames) #add separator
for (i in 1:length(featherNames)) {
  feather <- NULL
  feather <- import_bbsFeathers(newDir  = bbsDir,
                              filename = featherNames[i]) 
  feathers <- rbind(feathers, feather)
  rm(feather)
 
}
```


# Optional: Subset BBS data by taxonomic groupings, functional traits.
## Subset individual species
First, let's retrieve the species list from the BBS FTP server.
```{r getSpList, echo=TRUE}
spp <- get_speciesList()
glimpse(spp)
# This data frame contains common and scientific names, AOU numbers, taxonomic designations, and a taxonomic order (seq) for convenient sorting.
```

We can sort by AOU # (e.g. House Sparrow aou = 06882)
```{r subset_bySpeciesList, echo=TRUE}
subset_speciesList(myData = flBBS, aou.ind = 06882) %>% glimpse()
```

We could merge the bbs count data with the species list to avoid having to refer to AOU, then just subset using species name (e.g. 'House Sparrow).
```{r getHospData, echo=TRUE}
flBBS <- left_join(flBBS, spp)
flBBS %>% filter(commonName=="House Sparrow") %>% glimpse()
```

We can also  use the `subset_SpeciesList` as a convenient way to **remove**  taxonomic groups from the BBS data.
```{r , echo=TRUE, eval=FALSE}
subset_speciesList(flBBS, fam.ind = "Passeridae") 
subset_speciesList(flBBS, fam.ind = c("Passeridae", "Parulidae")) 
```



```{r , echo=TRUE}

```



```{r , echo=TRUE}

```


##################### END RUN #####################

## Next, we can subset the BBS data by species and/or functional traits (OPTIONAL but recommended)
Although subsetting the species is optional, I recommend removing waterfowl, wading birds, and shorebirds from analyses, especially as the spatial extent of the analysis increases. 

## Subset species according to AOU species codes (i.e. by family, genera, etc..)  
## Subset species according to functional traits (or body mass). 
### **Note: eval = F**
```{r subsetFxn, eval =F, message = T, warning = F}
# Load the functional trait and mass data, and munge/merge
funMass <-
    funcMass(dataWD = paste0(getwd(), "/data"),
             fxn = T, # get functional trait data?
             mass = F) # get body mass data?
# Combine the functional trqits and/or body mass
bbsData <-
    mergeFunMassBBS(bbsData = feather, funMass = funMass)
rm(funMass)